---
title: "Skills Development for Adolescent Girls"
subtitle: "Investment case, East Java, Indonesia"
author: "Hernando Grueso, PhD · UNICEF"
date: "7 October 2026"
lang: en
format:
  revealjs:
    theme: default
    slide-number: c/t
    transition: fade
    embed-resources: true
    fig-align: center
    include-in-header:
      text: |
        <style>
        .reveal .section-divider h1,
        .reveal .section-divider h2 { color:#fff !important; }
        .reveal em { color:#E2007A; font-style:normal; font-weight:600; }
        .reveal .placeholder {
        border:2px dashed #b9c2cc; border-radius:12px; background:#f6f8fa;
        padding:.7em 1em; color:#5a6b73; font-size:.8em; }
        .reveal .placeholder .ph-tag {
        font-family:monospace; color:#00377C; font-weight:700;
        text-transform:uppercase; font-size:.8em; letter-spacing:.04em; }
        .reveal .note { color:#cdd8e3; font-size:.7em; }
        .reveal table { font-size:.46em; }
        .reveal table th { color:#00377C; border-bottom:2px solid #d9e2ec; }
        .reveal .tight-table table { font-size:.74em; border-collapse:collapse; width:100%; }
        .reveal .tight-table table td, .reveal .tight-table table th { padding:2px 5px; }
        .reveal .big-table table { font-size:.85em; border-collapse:collapse; width:100%; }
        .reveal .big-table table td, .reveal .big-table table th { padding:4px 6px; }
        .vc-row { display:flex; align-items:center; justify-content:center; gap:.4em; flex-wrap:wrap; margin-top:1.2em; }
        .vc-box { background:#eaf6fc; border:1.5px solid #1CABE2; color:#00377C; border-radius:10px;
                  padding:.55em .8em; font-weight:600; font-size:.62em; text-align:center; min-width:8.5em; }
        .vc-box.benefit { background:#e8f5e9; border-color:#2e7d32; color:#1b5e20; }
        .vc-box.cost { background:#fde9f2; border-color:#E2007A; color:#a3005a; }
        .vc-arrow { font-size:1.3em; color:#9aa7ad; }
        /* compact horizontal value-chain flow, fits on one slide */
        .vc-flow { display:flex; align-items:center; justify-content:center; flex-wrap:nowrap;
                   gap:.25em; margin-top:1.6em; }
        .vc-node { background:#eaf6fc; border:1.5px solid #1CABE2; color:#00377C; border-radius:8px;
                   padding:.5em .55em; font-weight:700; font-size:.62em; text-align:center;
                   line-height:1.25; flex:1 1 0; }
        .vc-node small { display:block; font-weight:400; font-size:.85em; color:#3a6b85; margin-top:.2em; }
        .vc-flow-arrow { font-size:1.1em; color:#9aa7ad; flex:0 0 auto; }
        .vc-result-row { display:flex; align-items:center; justify-content:center; gap:.8em; margin-top:1.4em; }
        .vc-result-box { border-radius:10px; padding:.7em 1.2em; font-weight:700; font-size:.72em; text-align:center; }
        .vc-result-box.benefit { background:#e8f5e9; border:1.5px solid #2e7d32; color:#1b5e20; }
        .vc-result-box.cost { background:#fde9f2; border:1.5px solid #E2007A; color:#a3005a; }
        .vc-result-box.bcr { background:#fff7e6; border:1.5px solid #FFC20E; color:#7a5b00; }
        .bcr-grid { display:flex; flex-direction:column; gap:.9em; margin-top:.8em; }
        .bcr-col { border-radius:10px; padding:1em 1.3em; font-size:.78em; }
        .bcr-col.counted { background:#e8f5e9; border:1.5px solid #2e7d32; }
        .bcr-col.notcounted { background:#f4f4f4; border:1.5px solid #999; }
        .bcr-col h4 { margin-top:0; font-size:1.15em; }
        .toc-grid { font-size:.9em; line-height:1.6; margin-top:1.2em; }
        .toc-grid p { margin-bottom:.9em; }
        .toc-grid b { color:#00377C; }
        .kf-list { font-size:.9em; line-height:1.4; }
        .kf-list p { margin-bottom:.7em; }
        .reveal h2 { font-size:1.25em !important; line-height:1.15 !important; }
        .vmid { margin-top:2.2em; }
        .kf-list b { color:#E2007A; }
        .stat-row { display:flex; justify-content:center; gap:1em; margin-top:1em; }
        .stat { flex:1; text-align:center; background:#f6f8fa; border-radius:12px; padding:.8em .4em; }
        .stat .n { font-size:1.9em; font-weight:800; color:#E2007A; line-height:1.1; }
        .stat .l { font-size:.55em; color:#374649; margin-top:.4em; }
        .msg { font-size:.62em; color:#374649; margin-top:.4em; }
        </style>
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressPackageStartupMessages({ library(readr); library(dplyr); library(knitr); library(tibble) })
FIG <- function(file, width = "90%") {
  p <- file.path("..", "output", "figures", file)
  if (file.exists(p))
    cat(sprintf("![](%s){fig-align=\"center\" width=%s}\n", p, width))
  else
    cat("::: {.placeholder}\n<span class='ph-tag'>\u25a2 pending data</span><br>Figure pending: run the corresponding script (see README) to generate it.\n:::\n")
}
.rdc <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), col_types = readr::cols(.default = "c")), error = function(e) NULL)
.sc <- .rdc("projections/bcr_scenarios_table.csv")
BCR_C <- if (!is.null(.sc)) .sc$BCR[.sc$Scenario == "Central"] else "2.5"
BCR_FLOOR <- if (!is.null(.sc)) .sc$BCR[grepl("Retention channel only", .sc$Scenario)] else "1.0"
.mc <- .rdc("models/mincer_ifls_controls.csv")
MINCER_C <- if (!is.null(.mc)) { v <- .mc$`Return per year`[.mc$Sex == "Women" & grepl("^\\(4\\)", .mc$Model)]; sprintf("%.1f%%", 100 * as.numeric(v[1])) } else "9.6%"
.tn <- .rdc("projections/bcr_tornado.csv")
if (!is.null(.tn)) { .o <- .tn[!grepl("NEET|premium", .tn$Parameter), ]; .v <- as.numeric(c(.o$`BCR at low value`, .o$`BCR at high value`))
  RANGE_LO <- sprintf("%.2f", min(.v)); RANGE_HI <- sprintf("%.2f", max(.v)) } else { RANGE_LO <- "0.71"; RANGE_HI <- "1.57" }
TAB <- file.path("..", "output", "tables", "t8_target_funnel.csv")
```

## Contents

::: {.toc-grid}
**1 · The problem**: who is left out, where, and at what cost

**2 · Theory of change**: what the programme does and why it should work

**3 · Return on investment**: key assumptions and results

**4 · Key messages**

**Appendix**: regression results, methods and data
:::

# 1 · The problem {.section-divider background-color="#00377C"}

## Girls attend school at the same rate as boys, but their NEET rate is close to double {.smaller}
```{r f1a, results='asis'}
FIG("f1a_outcomes_by_sex_en.png", "78%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 15–24: 51% in school (boys 50%), 23% NEET (boys 13%), 25% employed (boys 37%). The gap opens at the transition out of school. SUSENAS Maret 2025, survey-weighted, 95% CIs.
:::

## East Java's NEET gender gap is wider than the national average {.smaller}
```{r f19, results='asis'}
FIG("f19_national_benchmark_en.png", "76%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls-minus-boys NEET gap: 10.3 points in East Java against 7.9 points nationally. Same survey, same definition, all provinces.
:::

## The gap costs an estimated USD 4.1 billion in forgone lifetime productivity {.smaller}
```{r f15, results='asis'}
FIG("f15_lifetime_loss_en.png", "80%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
About 288,000 girls are NEET in excess of the boys' rate. Each would be expected to earn about USD 14,100 more over a working lifetime (present value) if she had the boys' NEET rate.
:::

## Madura has higher NEET for girls and boys, with a similar gender gap {.smaller}
```{r f20, results='asis'}
FIG("f20_madura_en.png", "82%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls' NEET is 34% in Madura (Bangkalan, Sampang, Pamekasan, Sumenep) against 22% in the rest of East Java; boys' NEET is 23% against 12%. The gender gap is similar (11 vs 10 points), and fewer girls are in school (45% vs 52%). Madura is a priority area for a pilot.
:::

## Child marriage and outcomes: rest of East Java {.smaller}
```{r f24b, results='asis'}
FIG("f24_marriage_rest_en.png", "76%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 15–24 outside Madura, by whether they married before 18, survey-weighted (95% CIs). Regression tests by region are in the appendix. Early pregnancy before 18 overlaps almost entirely with child marriage in this data, so only child marriage is shown. Association, not causation.
:::

## Child marriage and outcomes: Madura {.smaller}
```{r f24a, results='asis'}
FIG("f24_marriage_madura_en.png", "76%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 15–24 in Bangkalan, Sampang, Pamekasan and Sumenep. Fewer girls, so wider intervals. Association, not causation.
:::

## Married girls who work are mostly in unpaid family work or own businesses {.smaller}
```{r f21, results='asis'}
FIG("f21_employment_type_en.png", "72%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Among employed girls married before 18, 35% are wage employees (others 75%) and 31% are unpaid family workers (others 12%); 22% work under 20 hours a week (others 12%). Their employment is lower-quality work. Based on 78 employed married girls, so indicative. SUSENAS has no earnings; the appendix shows earnings by type of work for young women from IFLS5.
:::

## Girls' NEET fell from 2024 to 2025, slightly more than in the rest of Indonesia {.smaller}
```{r f23, results='asis'}
FIG("f23_trend_ej_vs_national_en.png", "70%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls' NEET fell 5.6 points in East Java and 4.5 in the rest of Indonesia; child marriage fell 1.3 and 0.7 points. Tertiary enrolment at 18–20 rose 3.2 points, less than elsewhere (5.0). Most of the NEET decline is national, partly reflecting a change in the survey's activity question, so it cannot be attributed to a programme. A district-level comparison with programme districts is pending.
:::

## Senior-secondary enrolment also rose, in East Java and nationally {.smaller}
```{r f25, results='asis'}
FIG("f25_secondary_enrolment_trend_en.png", "70%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 16–18 enrolled in senior secondary, 2024 and 2025. A rise shared with the rest of Indonesia points to national drivers (such as PIP) rather than something specific to East Java.
:::

# 2 · Theory of change {.section-divider background-color="#00377C"}

## Completing senior secondary is associated with 20 points lower NEET for girls {.smaller}
```{r f22, results='asis'}
FIG("f22_completion_left_school_en.png", "74%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Among 18–20 year olds who have left school, girls who completed senior secondary have a NEET rate of 41%, against 62% for those who did not. Adjusted for age and area (girls 18–24), completion is associated with a 20-point lower NEET rate (p&lt;0.001). For boys there is no difference. Keeping girls in school to completion is the programme's main lever.
:::

## Evidence from Indonesia {.smaller}

::: {.tight-table}
| Programme | Design | Result | Relevance |
|---|---|---|---|
| **PKH** cash transfer (Priebe &amp; Sumarto, 2025) | Quasi-experimental | Girls' junior-secondary enrolment +5.2pp; child marriage −3.5pp | A cash component reduces child marriage |
| **PIP** education assistance (Ulfa &amp; Rezki, 2024) | Quasi-experimental | School attendance +11.4pp | Anchors the effect used in the ROI |
| **PIP**, girls' senior secondary (Caniago, 2021) | Quasi-experimental | Girls' senior-secondary enrolment +24.6pp | Targets the stage where girls drop out |
| **Vocational training** (Pritadrajati, 2022) | Quasi-experimental | Female LFPR +3.4pp; income +5.6% | Skills translate into work |
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
In SUSENAS 2025, girls in PKH households have lower child-marriage rates than other girls (3% vs 4–6%), even though PKH targets poorer households. Consistent with the PKH evidence, though not a causal estimate.
:::

## Theory of change

```{mermaid}
%%| fig-width: 13
flowchart LR
  T["Target: girls 15–18<br/>poorest households,<br/>Madura districts"]:::t
  A1["Cash+: link to PKH / PIP,<br/>mentoring and<br/>marriage-delay outreach"]:::a
  A2["Digital and 21st-century<br/>skills in Double Track"]:::a
  O1["Child marriage<br/>delayed"]:::o
  O2["Girls stay in school<br/>and complete<br/>senior secondary"]:::o
  O3["Skills for<br/>work"]:::o
  R["Transition to work<br/>or higher education:<br/>lower NEET"]:::r
  I["Higher lifetime<br/>earnings"]:::r
  T --> A1
  T --> A2
  A1 --> O1
  A1 --> O2
  A2 --> O3
  O1 --> O2
  O2 --> R
  O3 --> R
  R --> I
  classDef t fill:#f4f4f4,stroke:#999,color:#333
  classDef a fill:#eaf6fc,stroke:#1CABE2,color:#00377C
  classDef o fill:#fff7e6,stroke:#FFC20E,color:#7a5b00
  classDef r fill:#e8f5e9,stroke:#2e7d32,color:#1b5e20
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Each arrow is backed by evidence from Indonesia: cash transfers delay girls' marriage and raise their enrolment (PKH, girls); education assistance raises girls' senior-secondary enrolment (PIP, girls); completion is associated with 20 points lower NEET for girls (this analysis); skills raise women's employment and income (vocational training, women). Mentoring and outreach have only international evidence for girls (ELA, Room to Read). The cash component links girls to existing PKH and PIP rather than creating a new transfer.
:::

# 3 · Return on investment {.section-divider background-color="#00377C"}

## Key assumptions {.smaller}

::: {.big-table}
| Assumption | Value | Source |
|---|---|---|
| Reduction in girls' NEET | 5 points | PIP attendance effect (+11.4pp), 45% assumed to come from girls who would be NEET |
| Extra schooling for a girl kept in school | 3 years | Senior-secondary cycle |
| Earnings gain per year of schooling | `r MINCER_C` | IFLS5 wage return for women, controlling for region, parents' education and selection into paid work |
| Female earnings | Rp20 million/year | Below East Java minimum wage 2025, reflecting informal work |
| Female labour force participation | 52% | Sakernas, East Java |
| Earnings growth / discount rate | 2% / 3% per year | Career growth; World Bank education default |
| Skills earnings premium, all exposed girls who work | 5.6% for 10 years | Vocational training raised women's income by 5.6% (Pritadrajati 2022) |
| Programme cost | USD 200 per girl | USD 58 per year for 3 years, plus 15% overhead; Cash+ links to existing PKH/PIP |
:::

## How the return is built {.smaller}

::: {.vmid}
::: {.big-table}
```{r roi-steps}
st <- tryCatch(readr::read_csv(file.path("..","output","projections","bcr_steps.csv"), col_types = readr::cols(.default = "c")), error = function(e) NULL)
if (!is.null(st)) knitr::kable(st, align = "lr")
```
:::
:::

## Every USD 1 invested returns about USD `r BCR_C` {.smaller}

::: {.vmid}
```{r roi-stats, results='asis'}
if (!is.null(.sc) && exists("BCR_C")) {
  c0 <- .sc[.sc$Scenario == "Central", ]
  cat(sprintf('<div class="stat-row">
<div class="stat"><div class="n">USD %s</div><div class="l">return per USD 1, both channels</div></div>
<div class="stat"><div class="n">USD %s</div><div class="l">return per USD 1, retention channel only (floor)</div></div>
<div class="stat"><div class="n">USD %s</div><div class="l">cost per girl (3 years)</div></div>
<div class="stat"><div class="n">USD %s</div><div class="l">cost per girl moved out of NEET</div></div>
</div>', c0$BCR, BCR_FLOOR, c0$`Cost (USD/girl)`, c0$`Cost per NEET case averted (USD)`))
}
```
:::

::: {.kf-list style="font-size:.78em; margin-top:1em;"}
- **Retention channel**: 5 in 100 exposed girls stay in school and avoid NEET; each gains about three years of schooling and a lifetime of higher earnings.
- **Skills channel**: every exposed girl who works earns more, as vocational training has shown in Indonesia.
- Both channels value earnings only. Delayed marriage, health, intergenerational and fiscal benefits are excluded, so this is a lower bound.
:::

## The return depends mainly on the programme's effect on NEET {.smaller}

::: {.vmid}
::: {.tight-table}
```{r bcr-table}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
sc <- tryCatch(readr::read_csv(file.path("..","output","projections/bcr_scenarios_table.csv"), col_types = readr::cols(.default = "c")), error=function(e) NULL)
if (!is.null(sc)) knitr::kable(sc, align = "lrrrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
The floor is the retention channel alone; the central case adds the skills premium; the other rows show what happens at the 2025 minimum wage, and if a Cash+ stipend for the poorest 40% is added to the cost. With other assumptions varied one at a time, the central return stays between USD `r RANGE_LO` and USD `r RANGE_HI` (appendix).
:::

# 4 · Key messages {.section-divider background-color="#00377C"}

## Key messages {.smaller}

::: {.kf-list}
**1.** Girls in East Java stay in school as long as boys but are close to twice as likely to be NEET, a wider gap than nationally. The gap costs about USD 4.1 billion in forgone lifetime productivity.

**2.** Madura is the priority area: NEET is higher for both sexes, and fewer girls are in school.

**3.** Child marriage is strongly associated with leaving school and with lower-quality work; poorer girls face higher risk.

**4.** Completing senior secondary is associated with 20 points lower NEET for girls, which makes school completion the programme's main lever.

**5.** A package of skills, mentoring and links to PKH/PIP costing about USD 200 per girl returns about USD `r BCR_C` per USD 1 on earnings alone, and at least USD `r BCR_FLOOR` even if its only effect is keeping girls in school.
:::

# Appendix {.section-divider background-color="#00377C"}

## Regression results: child marriage and outcomes {.smaller}

::: {.vmid}
::: {.tight-table}
```{r marriage-tests}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
mt <- rd("models/marriage_tests.csv")
if (!is.null(mt)) knitr::kable(mt |> dplyr::select(Hypothesis, `Effect (pp)`, SE, p, N), align = "lrrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Linear-probability models, girls 15–24, survey-weighted, PSU-clustered. **Controls: age and urban/rural.** Poverty is not controlled for because the 2025 expenditure module is not yet available. The association with employment is small and not statistically significant. Association, not causation.
:::

## Child marriage and outcomes by region: regression results {.smaller}

::: {.vmid}
::: {.tight-table}
```{r marriage-region-tests}
mr <- tryCatch(readr::read_csv(file.path("..","output","models","followup2_A_marriage_tests_by_region.csv"), show_col_types = FALSE), error = function(e) NULL)
if (!is.null(mr)) knitr::kable(mr, align = "llrrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Linear-probability models, girls 15–24, survey-weighted, PSU-clustered; controls for age and urban/rural. Effects in percentage points.
:::

## Earnings by type of work, young women (IFLS5) {.smaller}

::: {.vmid}
::: {.tight-table}
```{r earnings-status}
es <- tryCatch(readr::read_csv(file.path("..","output","models","followup2_C_earnings_by_status_ifls.csv"), col_types = readr::cols(.default = "c")), error = function(e) NULL)
if (!is.null(es)) knitr::kable(es, align = "lrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Women 15–30, IFLS5 (2014/15 prices). SUSENAS collects no earnings. Unpaid family workers have no earnings by definition.
:::

## NEET among girls, by district {.smaller}
```{r f6, results='asis'}
FIG("f6_neet_by_kabkota_en.png", "60%")
```


## Digital access and use {.smaller}
```{r f2, results='asis'}
FIG("f2_digital_by_area_sex_en.png", "86%")
```


## Early pregnancy and outcomes {.smaller}
```{r f13, results='asis'}
FIG("f13_pregnancy_pathway_en.png", "78%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Recorded only for ever-married women, so a lower bound. Girls with a first pregnancy before 18 are almost all girls married before 18, so the main deck reports child marriage only.
:::

## Child marriage and NEET by household PKH receipt {.smaller}

::: {.tight-table}
```{r pkh}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
pk <- rd("models/followup_D_pkh.csv")
if (!is.null(pk)) knitr::kable(pk, align = "llrrr")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 15–24, SUSENAS 2025. PKH targets poorer households, so lower rates among recipients are consistent with a protective effect, but are not a causal estimate.
:::

## Cost components {.smaller}

::: {.big-table}
```{r costing-table}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
ct <- rd("projections/bcr_costing_table.csv")
if (!is.null(ct)) knitr::kable(ct, align = "lrl")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
These are benchmark unit costs from comparable programmes, not quotes from Double Track; they are placeholders until the provincial government shares actual budgets. The transfer itself is not costed because the Cash+ component links girls to existing PKH and PIP entitlements; a new stipend for the poorest 40% at the PIP senior-secondary rate would add about USD 46 per girl per year (ROI table).
:::

## Sensitivity of the return to each assumption {.smaller}

::: {.tight-table}
```{r tornado}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
tn <- rd("projections/bcr_tornado.csv")
if (!is.null(tn)) knitr::kable(tn |> dplyr::select(-spread), align = "lrr")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Each assumption varied on its own, others held at central values.
:::

## What the return counts {.smaller}

::: {.bcr-grid}
<div class="bcr-col counted">
<h4>Counted</h4>
Higher schooling leads to a lifetime earnings premium for girls moved out of NEET, using the female return to schooling over a discounted working life.
</div>
<div class="bcr-col notcounted">
<h4>Excluded <span style="font-weight:400;">(so the return is a lower bound)</span></h4>
Delayed child marriage · health and productivity spillovers · intergenerational gains · fiscal returns · broader empowerment
</div>
:::

## Comparable international evidence {.smaller}

```{r evidence-table}
tibble::tribble(
  ~Programme, ~Design, ~"Girls / age", ~Focus, ~"Key result", ~"Fit for East Java",
  "Adolescent girls' wellbeing, Bergström &amp; Özler 2021", "Review (104 studies)", "Adolescent girls", "Multi-sector interventions", "Integrated packages beat single-sector ones", "Synthesis backbone for the ToC",
  "NEET in Indonesia, Faizah, Isbah &amp; Azca 2025", "Qualitative", "Youth (gendered)", "Drivers of girls' NEET", "Non-education factors (norms, care, marriage) drive girls' NEET", "Indonesia-specific NEET context",
  "ELA girls' clubs, Bandiera 2018 (Sierra Leone)", "RCT", "Girls 12–25", "Vocational + life skills", "Enrolment +8.5pp; more income-generating activity", "Closest Skills4Girls-in-Double-Track model",
  "Female digital-employment bootcamp, Alzúa 2020 (Argentina/Colombia)", "RCT", "Women 18–50", "Coding for employment", "Large coding-skill gains", "Core digital-skills → jobs arm",
  "Irumi, Näslund-Hadley 2025 (Paraguay)", "RCT", "Pupils", "Computational thinking", "+0.09 SD CT", "Structured CT pedagogy",
  "Robotics &amp; Coding, Gülen 2025 (Turkey)", "RCT", "Girls 14–16", "Robotics / coding", "Higher STEM-career orientation", "Girls' STEM, secondary age",
  "Samata, Prakash 2019 (India)", "RCT", "Poor girls 13–16", "Secondary-transition support", "More girls transition; norm shift", "South Asia; senior-secondary transition",
  "Room to Read GEP, Edmonds 2021 (India)", "RCT", "Girls", "Life skills + mentoring", "Dropout −3.5 to −4.1pp", "Mentoring/pedagogy layer cuts dropout",
  "Negotiating a Better Future, Ashraf 2025 (Zambia)", "RCT", "Girls ~14", "Negotiation skills", "+0.26 years schooling (10-yr)", "Durable, cheap soft-skills layer",
  "OLPC / Plan Ceibal, de Melo 2014 (Uruguay)", "Quasi / review", "Pupils", "Laptops (hardware-led)", "Learning ≈ null (−0.05 SD)", "Caution: hardware alone fails → fund pedagogy"
) |> kable(align = "llllll")
```

## Returns to schooling: specification and results {.smaller}

::: {style="font-size:0.8em"}
**Wage-based** (IFLS5 2014/15, adults 20–60, estimated separately for women and men, weighted):

$$\ln(w_i) = \beta_0 + \beta_1 S_i + \beta_2 \text{Exp}_i + \beta_3 \text{Exp}_i^2 + \beta_4 \text{Urban}_i + \gamma_p + \phi_f + \phi_m + \delta \lambda_i + \varepsilon_i$$

$w$: monthly wage; $S$: years of schooling; Exp: potential experience (age minus schooling minus 6); $\gamma_p$: province fixed effects; $\phi_f$, $\phi_m$: father's and mother's education (fixed effects); $\lambda$: selection term from a first-stage model of working, in which children under 5 in the household predict whether a woman works but not her wage (Heckman two-step).
:::

::: {.tight-table}
```{r mincer-controls}
mc <- tryCatch(readr::read_csv(file.path("..","output","models","mincer_ifls_controls.csv"), show_col_types = FALSE), error = function(e) NULL)
if (!is.null(mc)) knitr::kable(mc |> dplyr::mutate(`Return per year` = sprintf("%.1f%%", 100 * `Return per year`),
                                                   SE = sprintf("%.1f", 100 * SE)) |>
                                 dplyr::select(Sex, Model, `Return per year`, SE, N), align = "llrrr")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
The ROI uses the women's estimate with all controls (model 4, `r MINCER_C`). The consumption-based return from SUSENAS 2024 (5.6% for women; household per-capita expenditure regressed on schooling, schooling², age, age² and urban) is the low case in the sensitivity analysis.
:::

## Data sources {.smaller}

::: {style="font-size:0.95em"}
- **SUSENAS Maret 2025** (BPS): NEET, school attendance and completion, child marriage, early pregnancy, type of work and hours, PKH receipt and digital access, for East Java and all provinces. East Java: 98,738 individuals, 12,599 aged 15–24, of whom 6,060 girls.
- **SUSENAS Maret 2024** (BPS): the 2024 comparison, and per-capita expenditure for the consumption-based return to schooling.
- **IFLS5 2014/15**: wages for the wage-based return to schooling.
- **Sakernas** (BPS): female labour force participation in East Java.
- **UMP Jawa Timur 2025**: benchmark for female earnings.
- **Published evaluations** of PKH, PIP and vocational training: the programme effect used in the return on investment.

All SUSENAS estimates are survey-weighted, with PSU/strata-clustered standard errors.
:::

## Definitions {.smaller}

::: {style="font-size:0.95em"}
- **NEET (15–24)**: not attending school (`R611`) and not working (`R703_A`, `R704`).
- **Married before 18**: age at first marriage (`R409`) under 18.
- **Senior-secondary completion**: highest certificate (`R615`) at SMA/SMK/MA level or above.
- **Type of work**: status in main job (`R706`); hours (`R707`).
- **Madura**: Bangkalan, Sampang, Pamekasan, Sumenep.

The 2025 questionnaire replaced the single main-activity question with a multi-select block, so year-on-year changes partly reflect the new question; comparisons with the rest of Indonesia use the same definition in both places.
:::
