---
title: "Skills Development for Adolescent Girls"
subtitle: "Investment case · East Java, Indonesia"
author: "Hernando Grueso, PhD · UNICEF"
date: "31 August 2026"
lang: en
format:
  revealjs:
    theme: default
    slide-number: c/t
    transition: fade
    embed-resources: true
    fig-align: center
    include-in-header:
      text: |
        <style>
        .reveal .section-divider h1,
        .reveal .section-divider h2 { color:#fff !important; }
        .reveal em { color:#E2007A; font-style:normal; font-weight:600; }
        .reveal .placeholder {
        border:2px dashed #b9c2cc; border-radius:12px; background:#f6f8fa;
        padding:.7em 1em; color:#5a6b73; font-size:.8em; }
        .reveal .placeholder .ph-tag {
        font-family:monospace; color:#00377C; font-weight:700;
        text-transform:uppercase; font-size:.8em; letter-spacing:.04em; }
        .reveal .note { color:#cdd8e3; font-size:.7em; }
        .reveal table { font-size:.46em; }
        .reveal table th { color:#00377C; border-bottom:2px solid #d9e2ec; }
        .reveal .tight-table table { font-size:.74em; border-collapse:collapse; width:100%; }
        .reveal .tight-table table td, .reveal .tight-table table th { padding:2px 5px; }
        .reveal .big-table table { font-size:.85em; border-collapse:collapse; width:100%; }
        .reveal .big-table table td, .reveal .big-table table th { padding:4px 6px; }
        .vc-row { display:flex; align-items:center; justify-content:center; gap:.4em; flex-wrap:wrap; margin-top:1.2em; }
        .vc-box { background:#eaf6fc; border:1.5px solid #1CABE2; color:#00377C; border-radius:10px;
                  padding:.55em .8em; font-weight:600; font-size:.62em; text-align:center; min-width:8.5em; }
        .vc-box.benefit { background:#e8f5e9; border-color:#2e7d32; color:#1b5e20; }
        .vc-box.cost { background:#fde9f2; border-color:#E2007A; color:#a3005a; }
        .vc-arrow { font-size:1.3em; color:#9aa7ad; }
        /* compact horizontal value-chain flow, fits on one slide */
        .vc-flow { display:flex; align-items:center; justify-content:center; flex-wrap:nowrap;
                   gap:.25em; margin-top:1.6em; }
        .vc-node { background:#eaf6fc; border:1.5px solid #1CABE2; color:#00377C; border-radius:8px;
                   padding:.5em .55em; font-weight:700; font-size:.62em; text-align:center;
                   line-height:1.25; flex:1 1 0; }
        .vc-node small { display:block; font-weight:400; font-size:.85em; color:#3a6b85; margin-top:.2em; }
        .vc-flow-arrow { font-size:1.1em; color:#9aa7ad; flex:0 0 auto; }
        .vc-result-row { display:flex; align-items:center; justify-content:center; gap:.8em; margin-top:1.4em; }
        .vc-result-box { border-radius:10px; padding:.7em 1.2em; font-weight:700; font-size:.72em; text-align:center; }
        .vc-result-box.benefit { background:#e8f5e9; border:1.5px solid #2e7d32; color:#1b5e20; }
        .vc-result-box.cost { background:#fde9f2; border:1.5px solid #E2007A; color:#a3005a; }
        .vc-result-box.bcr { background:#fff7e6; border:1.5px solid #FFC20E; color:#7a5b00; }
        .bcr-grid { display:flex; flex-direction:column; gap:.9em; margin-top:.8em; }
        .bcr-col { border-radius:10px; padding:1em 1.3em; font-size:.78em; }
        .bcr-col.counted { background:#e8f5e9; border:1.5px solid #2e7d32; }
        .bcr-col.notcounted { background:#f4f4f4; border:1.5px solid #999; }
        .bcr-col h4 { margin-top:0; font-size:1.15em; }
        .toc-grid { font-size:.9em; line-height:1.6; margin-top:1.2em; }
        .toc-grid p { margin-bottom:.9em; }
        .toc-grid b { color:#00377C; }
        .kf-list { font-size:.68em; line-height:1.85; }
        .kf-list b { color:#E2007A; }
        </style>
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressPackageStartupMessages({ library(readr); library(dplyr); library(knitr); library(tibble) })
FIG <- function(file, width = "90%") {
  p <- file.path("..", "output", "figures", file)
  if (file.exists(p))
    cat(sprintf("![](%s){fig-align=\"center\" width=%s}\n", p, width))
  else
    cat("::: {.placeholder}\n<span class='ph-tag'>\u25a2 pending data</span><br>Figure pending — run the corresponding script (see README) to generate it.\n:::\n")
}
TAB <- file.path("..", "output", "tables", "t8_target_funnel.csv")
```

## Contents

::: {.toc-grid}
**1 · Diagnostic** — the size and shape of the gap

**2 · Testing the relationships** — regression analysis of what is associated with NEET

**3 · Evidence &amp; Theory of Change** — what has worked in Indonesia, and the proposed programme

**4 · Costing &amp; ROI** — component costs and benefit-cost ratio

**5 · Key Findings**

**Appendix** — data, definitions, and methods
:::

# 1 · Diagnostic {.section-divider background-color="#00377C"}

## Girls, boys, and the school-to-work transition {.smaller}
```{r f1a, results='asis'}
FIG("f1a_outcomes_by_sex_en.png", "80%")
```

::: {style="font-size:0.6em; color:#7a7a7a;"}
Girls 15–24 are slightly ahead of boys in school attendance (51% vs 50%), but their NEET rate is close to double (23% vs 13%) and their employment rate is 12 points lower (25% vs 37%). The difference appears at the transition out of school, not within it. SUSENAS Maret 2025, survey-weighted; 95% CIs.
:::

## Forgone lifetime productivity from the NEET gap {.smaller}

```{r f15, results='asis'}
FIG("f15_lifetime_loss_en.png", "84%")
```

::: {style="font-size:0.6em; color:#7a7a7a;"}
**What this is**: the number of girls who are NEET in excess of the boys' rate, multiplied by the earnings each would be expected to make over a working lifetime, discounted to present value. Full calculation in `output/models/lifetime_loss.csv`.
:::

## The gap in national context {.smaller}

```{r f19, results='asis'}
FIG("f19_national_benchmark_en.png", "80%")
```

::: {style="font-size:0.6em; color:#7a7a7a;"}
The same NEET gender gap computed for every Indonesian province from the same survey, showing where East Java sits relative to the national average. See `output/models/national_benchmark.csv`.
:::

## NEET among girls, by district {.smaller}
```{r f6, results='asis'}
FIG("f6_neet_by_kabkota_en.png", "62%")
```

::: {style="font-size:0.55em; color:#7a7a7a;"}
SUSENAS Maret 2025, girls 15–24, survey-weighted; East Java districts with ≥15 sampled girls.
:::

## Digital access and use {.smaller}
```{r f2, results='asis'}
FIG("f2_digital_by_area_sex_en.png", "88%")
```

## Child marriage, schooling, and labour-market outcomes {.smaller}
```{r f9, results='asis'}
FIG("f9_marriage_pathway_en.png", "78%")
```

::: {style="font-size:0.56em; color:#7a7a7a;"}
Girls 15–24, survey-weighted (95% CIs). Girls who married before 18 have a NEET rate of 59% against 22% for those who did not, and 8% are still in school against 53%. Around 4% of East-Java girls 15–24 married before 18. **Association, not causation** — the direction is not identified, and most NEET girls did not marry before 18.
:::

## Early pregnancy, schooling, and labour-market outcomes {.smaller}
```{r f13, results='asis'}
FIG("f13_pregnancy_pathway_en.png", "84%")
```

::: {style="font-size:0.6em; color:#7a7a7a;"}
First pregnancy before 18 (SUSENAS Blok XV). SUSENAS records pregnancy history only for ever-married women, so these rates are a lower bound on the true prevalence.
:::

## Change between 2024 and 2025 {.smaller}

```{r f18, results='asis'}
FIG("f18_trend_2024_2025_en.png", "78%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Two consecutive SUSENAS rounds for girls 15–24. Girls' NEET fell 4.6 points and child marriage 1.3 points; school attendance and employment were essentially flat.
:::

# 2 · Testing the relationships {.section-divider background-color="#00377C"}

## Senior-secondary completion and NEET {.smaller}

```{r f16, results='asis'}
FIG("f16_completion_neet_en.png", "80%")
```

::: {style="font-size:0.56em; color:#7a7a7a;"}
NEET rates by senior-secondary completion status, girls and boys separately. Regression-adjusted effect sizes in `output/models/completion_neet.csv`. Association, not causation: girls who complete senior secondary may differ from those who do not in ways this data does not capture.
:::

## Regression results — marriage, pregnancy and exclusion {.smaller}
::: {.tight-table}
```{r marriage-tests}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output","models",p),
        show_col_types=FALSE), error=function(e) NULL)
mtab <- dplyr::bind_rows(rd("marriage_tests.csv"), rd("pregnancy_tests.csv")) |>
  dplyr::select(Hypothesis, `Effect (pp)`, SE, p)
if (nrow(mtab) > 0) knitr::kable(mtab, align="lrrr")
```
:::

::: {style="font-size:0.52em; color:#7a7a7a;"}
Linear-probability estimates, girls 15–24, survey-weighted, PSU-clustered. Effects in percentage points. Controls: age and urban/rural throughout, plus poverty where expenditure data is available. Robust to adding district fixed effects. Association, not causal identification.
:::

# 3 · Evidence &amp; Theory of Change {.section-divider background-color="#00377C"}

## Evidence from Indonesia {.smaller}

::: {.tight-table}
| Programme | Design | Result | Relevance |
|---|---|---|---|
| **PKH** conditional cash transfer (Priebe &amp; Sumarto, 2025) | Quasi-experimental | Girls' junior-secondary enrolment +5.2pp; child marriage −3.5pp | Direct evidence a cash component reduces child marriage in Indonesia |
| **PIP** education assistance (Ulfa &amp; Rezki, 2024) | Quasi-experimental | School attendance +11.4pp | Anchors the attendance effect used in the ROI model |
| **PIP** — girls' senior secondary (Caniago, 2021) | Quasi-experimental | Girls' senior-secondary enrolment +24.6pp | Senior-secondary transition, the stage this programme targets |
| **Vocational training** (Pritadrajati, 2022) | Quasi-experimental | Female LFPR +3.4pp; income +5.6%; BCR up to 22.4 | Indonesian precedent for skills-to-work returns |
| NEET in Indonesia (Faizah, Isbah &amp; Azca, 2025) | Qualitative | Norms, care duties and marriage drive girls' NEET | Explains the transition gap seen in Section 1 |
:::

::: {style="font-size:0.55em; color:#7a7a7a;"}
The ROI model in Section 4 is anchored on **PIP/PKH Indonesian effect sizes directly** — no cross-country transport discount is applied, because domestic evidence exists for this intervention type.
:::

## Comparable international evidence {.smaller}

```{r evidence-table}
tibble::tribble(
  ~Programme, ~Design, ~"Girls / age", ~Focus, ~"Key result", ~"Fit for East Java",
  "Adolescent girls' wellbeing — Bergström &amp; Özler 2021", "Review (104 studies)", "Adolescent girls", "Multi-sector interventions", "Integrated packages beat single-sector ones", "Synthesis backbone for the ToC",
  "NEET in Indonesia — Faizah, Isbah &amp; Azca 2025", "Qualitative", "Youth (gendered)", "Drivers of girls' NEET", "Non-education factors (norms, care, marriage) drive girls' NEET", "Indonesia-specific NEET context",
  "ELA girls' clubs — Bandiera 2018 (Sierra Leone)", "RCT", "Girls 12–25", "Vocational + life skills", "Enrolment +8.5pp; more income-generating activity", "Closest Skills4Girls-in-Double-Track model",
  "Female digital-employment bootcamp — Alzúa 2020 (Argentina/Colombia)", "RCT", "Women 18–50", "Coding for employment", "Large coding-skill gains", "Core digital-skills → jobs arm",
  "Irumi — Näslund-Hadley 2025 (Paraguay)", "RCT", "Pupils", "Computational thinking", "+0.09 SD CT", "Structured CT pedagogy",
  "Robotics &amp; Coding — Gülen 2025 (Turkey)", "RCT", "Girls 14–16", "Robotics / coding", "Higher STEM-career orientation", "Girls' STEM, secondary age",
  "Samata — Prakash 2019 (India)", "RCT", "Poor girls 13–16", "Secondary-transition support", "More girls transition; norm shift", "South Asia; senior-secondary transition",
  "Room to Read GEP — Edmonds 2021 (India)", "RCT", "Girls", "Life skills + mentoring", "Dropout −3.5 to −4.1pp", "Mentoring/pedagogy layer cuts dropout",
  "Negotiating a Better Future — Ashraf 2025 (Zambia)", "RCT", "Girls ~14", "Negotiation skills", "+0.26 years schooling (10-yr)", "Durable, cheap soft-skills layer",
  "OLPC / Plan Ceibal — de Melo 2014 (Uruguay)", "Quasi / review", "Pupils", "Laptops (hardware-led)", "Learning ≈ null (−0.05 SD)", "Caution: hardware alone fails → fund pedagogy"
) |> kable(align = "llllll")
```

::: {style="font-size:0.55em; color:#7a7a7a;"}
Effects as reported in each source. Used to inform programme design; the ROI effect size itself comes from the Indonesian evidence on the previous slide.
:::

## Theory of change

```{mermaid}
%%| fig-width: 13
flowchart LR
  A1["Digital &amp; 21st-C skills<br/>content + connectivity"]:::a
  A2["Girls-responsive pedagogy<br/>teacher training + mentoring"]:::b
  A3["Ancillary supports<br/>marriage-delay, flexible<br/>learning, safe spaces"]:::c
  O1["Skills acquired<br/>digital · STEM · socio-emotional"]:::o
  O2["Stay enrolled,<br/>complete senior secondary"]:::o
  M1["Transition to<br/>higher ed / work"]:::o
  R["Lower NEET ·<br/>narrower gender gap"]:::r
  I["Higher lifetime earnings ·<br/>productivity → 2045 Vision"]:::r
  A1 --> O1
  A2 --> O1
  A2 --> O2
  A3 --> O2
  O1 --> M1
  O2 --> M1
  M1 --> R --> I
  classDef a fill:#eaf6fc,stroke:#1CABE2,color:#00377C
  classDef b fill:#fde9f2,stroke:#E2007A,color:#a3005a
  classDef c fill:#f4f4f4,stroke:#999,color:#555
  classDef o fill:#fff7e6,stroke:#FFC20E,color:#7a5b00
  classDef r fill:#e8f5e9,stroke:#2e7d32,color:#1b5e20
```

::: {style="font-size:0.6em; color:#7a7a7a;"}
Embedded in Double Track / <em>Tambahan Keterampilan</em>. Measured against the ToR: cost per NEET averted, transition to higher education or work, BCR.
:::

# 4 · Costing &amp; ROI {.section-divider background-color="#00377C"}

## Component costs {.smaller}

All figures are **USD per girl, per year of programme exposure**.

::: {.big-table}
```{r costing-table}
suppressPackageStartupMessages(library(knitr))
ct <- tryCatch(readr::read_csv(file.path("..","output","projections","bcr_costing_table.csv"),
        show_col_types=FALSE), error=function(e) NULL)
if (!is.null(ct)) ct |> dplyr::select(-Source) |>
  kable(align="lrrr", caption="USD per girl per year — Low / Central / High")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
**Source/method:** Programme- and literature-benchmarked unit costs, using Indonesia-specific sources where available. Training and mentoring costs are annualised from comparable programmes; connectivity assumes shared labs rather than 1:1 devices. Ancillary activities reflect ToR guidance, and overheads follow UNICEF's 10–20% range. Low/central/high estimates capture sourcing uncertainty. Replace with Double Track actuals once provincial financing data are available.
:::

## Benefit value chain {.smaller}

::: {.vc-flow}
<div class="vc-node">NEET reduction (pp)<small>from PIP/PKH evidence</small></div>
<div class="vc-flow-arrow">→</div>
<div class="vc-node">Extra years of<br>senior-secondary schooling</div>
<div class="vc-flow-arrow">→</div>
<div class="vc-node">Female Mincer return<small>IFLS5 wage-based</small></div>
<div class="vc-flow-arrow">→</div>
<div class="vc-node">× wage<br>× LFP<br>× working life</div>
:::

::: {.vc-result-row}
<div class="vc-result-box benefit">Benefit, PV per girl</div>
<div class="vc-flow-arrow">÷</div>
<div class="vc-result-box cost">Cost = Σ components × exposure × (1+overhead)</div>
<div class="vc-flow-arrow">=</div>
<div class="vc-result-box bcr">BCR</div>
:::

::: {style="font-size:0.58em; color:#7a7a7a; margin-top:1.2em;"}
The effect chain runs left to right into the benefit; cost is computed independently, and the two combine into the BCR.
:::

## BCR and cost per NEET case averted {.smaller}

"NEET averted" is a count of girls, not a percentage point: the pp effect multiplied by girls exposed gives an expected number of cases prevented.

::: {.tight-table}
```{r bcr-table}
comma_fix <- function(x) {
  num <- suppressWarnings(as.numeric(gsub(",", "", x)))
  ifelse(!is.na(num), formatC(num, format="d", big.mark=","), x)
}
br <- tryCatch(readr::read_csv(file.path("..","output","projections","bcr_scenarios_table.csv"),
        show_col_types=FALSE), error=function(e) NULL)
if (!is.null(br)) br |>
  dplyr::select(Scenario, `Years edu gained`, `Benefit PV (USD/girl)`,
                `Cost (USD/girl)`, `NEET cases averted per exposed girl`, BCR,
                `Cost per full NEET case averted (USD)`) |>
  dplyr::mutate(`Cost per full NEET case averted (USD)` = comma_fix(`Cost per full NEET case averted (USD)`)) |>
  kable(align="lrrrrrr")
```
:::

::: {style="font-size:0.55em; color:#7a7a7a;"}
The Low scenario carries the null effect (its confidence interval includes zero), so its BCR is not defined — the return is conditional on a real effect, to be confirmed at pilot.
:::

## What the BCR counts {.smaller}

::: {.bcr-grid}
<div class="bcr-col counted">
<h4>Counted</h4>
Higher schooling → lifetime earnings premium for girls moved out of NEET, via the female Mincer return applied over the discounted working life.
</div>
<div class="bcr-col notcounted">
<h4>Excluded <span style="font-weight:400;">(so the BCR is a lower bound)</span></h4>
Intergenerational gains · delayed child marriage · health and productivity spillovers · fiscal returns · broader empowerment
</div>
:::

# 5 · Key Findings {.section-divider background-color="#00377C"}

## Key findings {.smaller}

::: {.kf-list}
**1.** Girls 15–24 match boys in school attendance but have close to double the NEET rate (23% vs 13%) and 12 points lower employment. The gap is at the school-to-work transition.

**2.** That gap carries a measurable present cost in forgone lifetime productivity, quantified in Section 1 and benchmarked against the national average.

**3.** Senior-secondary completion is associated with substantially lower NEET for both sexes — the mechanism the programme targets.

**4.** Child marriage and early pregnancy are each strongly associated with higher NEET and lower school attendance, though they account for a minority of NEET cases.

**5.** Girls' NEET fell 4.6 points and child marriage 1.3 points between 2024 and 2025, while attendance and employment held flat.

**6.** Indonesian evidence (PKH, PIP, vocational training) supports both the skills component and a cash-transfer component, and anchors the ROI model without cross-country extrapolation.

**7.** Open items pending external data: disability (2023 SUSENAS), Sakernas wage microdata, higher-education field of study (PDDikti), and the 2025 expenditure module for poverty analysis.
:::

# Appendix {.section-divider background-color="#00377C"}

## Data and sample {.smaller}

::: {style="font-size:0.84em"}
**Source.** SUSENAS Maret 2025, East Java (province 35) — 98,738 individuals; adolescents 15–24: **12,599** (6,060 girls). Survey-weighted, PSU/strata-clustered (`WEIND`, `PSU`, `STRATA`). SUSENAS Maret 2024 retained for the year-on-year comparison.

**Definitions:**

- **NEET (15–24)** = not currently attending school (`R611`) and not working — built from the activity block (`R703_A` worked, `R703_B` schooled, `R704` temporarily absent from a job).
- **Sex** `R405` · **age** `R407` · **married before 18** from age at first marriage (`R409`) · **early pregnancy** from age at first pregnancy (Blok XV, ever-married women only).
- **Educational attainment** from highest certificate held (`R615`).

**Note on variable numbering:** the 2025 questionnaire renumbered several variables relative to 2024 (school participation R610→R611, certificate R614→R615, internet R805→R808, pregnancy block 1501→1401) and replaced the single main-activity question with a multi-select block. The 2024 and 2025 cleaning scripts are therefore separate files.
:::

## Data availability {.smaller}

::: {.tight-table}
| Indicator | Status |
|---|---|
| NEET, schooling, employment, digital access | Available (2024 and 2025) |
| Child marriage, early pregnancy | Available; pregnancy is ever-married women only (lower bound) |
| Expenditure, wealth quintiles, official poverty | **Not in the 2025 merged file** — the consumption module is a separate BPS release, not yet obtained |
| Wages | Not collected in SUSENAS; IFLS5 used for the Mincer return |
| Disability | Not in the 2024 or 2025 KOR extract; available in SUSENAS 2023 |
| Higher-education field of study | Not collected in SUSENAS; PDDikti is the source |
| Programme costs | Pending from the provincial government |
:::

## The Mincer specification {.smaller}

Returns to schooling are estimated as:

$$\ln(Y_i) = \beta_0 + \beta_1 S_i + \beta_2 S_i^2 + \beta_3 \text{Age}_i + \beta_4 \text{Age}_i^2 + \beta_5 \text{Urban}_i + \varepsilon_i$$

- $Y_i$ — monthly wage (IFLS5 earners 20–60)
- $S_i$ — years of schooling, from highest certificate attained
- $\beta_1$ — the return per additional year of schooling
- $\beta_2$ — the quadratic term; positive and significant indicates returns that rise with schooling level
- Estimated weighted, with robust standard errors

::: {style="font-size:0.55em; color:#7a7a7a;"}
A consumption-based version of the same specification (using SUSENAS per-capita expenditure in place of wages) is available once the 2025 expenditure module is obtained; the 2024 estimate was +5.6%/yr for women, +5.2%/yr for men, with significant convexity.
:::
