---
title: "Skills Development for Adolescent Girls"
subtitle: "Investment case, East Java, Indonesia"
author: "Hernando Grueso, PhD · UNICEF"
date: "7 October 2026"
lang: en
format:
  revealjs:
    theme: default
    slide-number: c/t
    transition: fade
    embed-resources: true
    fig-align: center
    include-in-header:
      text: |
        <style>
        .reveal .section-divider h1,
        .reveal .section-divider h2 { color:#fff !important; }
        .reveal em { color:#E2007A; font-style:normal; font-weight:600; }
        .reveal .placeholder {
        border:2px dashed #b9c2cc; border-radius:12px; background:#f6f8fa;
        padding:.7em 1em; color:#5a6b73; font-size:.8em; }
        .reveal .placeholder .ph-tag {
        font-family:monospace; color:#00377C; font-weight:700;
        text-transform:uppercase; font-size:.8em; letter-spacing:.04em; }
        .reveal .note { color:#cdd8e3; font-size:.7em; }
        .reveal table { font-size:.46em; }
        .reveal table th { color:#00377C; border-bottom:2px solid #d9e2ec; }
        .reveal .tight-table table { font-size:.74em; border-collapse:collapse; width:100%; }
        .reveal .tight-table table td, .reveal .tight-table table th { padding:2px 5px; }
        .reveal .big-table table { font-size:.85em; border-collapse:collapse; width:100%; }
        .reveal .big-table table td, .reveal .big-table table th { padding:4px 6px; }
        .vc-row { display:flex; align-items:center; justify-content:center; gap:.4em; flex-wrap:wrap; margin-top:1.2em; }
        .vc-box { background:#eaf6fc; border:1.5px solid #1CABE2; color:#00377C; border-radius:10px;
                  padding:.55em .8em; font-weight:600; font-size:.62em; text-align:center; min-width:8.5em; }
        .vc-box.benefit { background:#e8f5e9; border-color:#2e7d32; color:#1b5e20; }
        .vc-box.cost { background:#fde9f2; border-color:#E2007A; color:#a3005a; }
        .vc-arrow { font-size:1.3em; color:#9aa7ad; }
        /* compact horizontal value-chain flow, fits on one slide */
        .vc-flow { display:flex; align-items:center; justify-content:center; flex-wrap:nowrap;
                   gap:.25em; margin-top:1.6em; }
        .vc-node { background:#eaf6fc; border:1.5px solid #1CABE2; color:#00377C; border-radius:8px;
                   padding:.5em .55em; font-weight:700; font-size:.62em; text-align:center;
                   line-height:1.25; flex:1 1 0; }
        .vc-node small { display:block; font-weight:400; font-size:.85em; color:#3a6b85; margin-top:.2em; }
        .vc-flow-arrow { font-size:1.1em; color:#9aa7ad; flex:0 0 auto; }
        .vc-result-row { display:flex; align-items:center; justify-content:center; gap:.8em; margin-top:1.4em; }
        .vc-result-box { border-radius:10px; padding:.7em 1.2em; font-weight:700; font-size:.72em; text-align:center; }
        .vc-result-box.benefit { background:#e8f5e9; border:1.5px solid #2e7d32; color:#1b5e20; }
        .vc-result-box.cost { background:#fde9f2; border:1.5px solid #E2007A; color:#a3005a; }
        .vc-result-box.bcr { background:#fff7e6; border:1.5px solid #FFC20E; color:#7a5b00; }
        .bcr-grid { display:flex; flex-direction:column; gap:.9em; margin-top:.8em; }
        .bcr-col { border-radius:10px; padding:1em 1.3em; font-size:.78em; }
        .bcr-col.counted { background:#e8f5e9; border:1.5px solid #2e7d32; }
        .bcr-col.notcounted { background:#f4f4f4; border:1.5px solid #999; }
        .bcr-col h4 { margin-top:0; font-size:1.15em; }
        .toc-grid { font-size:.9em; line-height:1.6; margin-top:1.2em; }
        .toc-grid p { margin-bottom:.9em; }
        .toc-grid b { color:#00377C; }
        .kf-list { font-size:.9em; line-height:1.4; }
        .kf-list p { margin-bottom:.7em; }
        .reveal h2 { font-size:1.25em !important; line-height:1.15 !important; }
        .vmid { margin-top:2.2em; }
        .kf-list b { color:#E2007A; }
        .stat-row { display:flex; justify-content:center; gap:1em; margin-top:1em; }
        .stat { flex:1; text-align:center; background:#f6f8fa; border-radius:12px; padding:.8em .4em; }
        .stat .n { font-size:1.9em; font-weight:800; color:#E2007A; line-height:1.1; }
        .stat .l { font-size:.55em; color:#374649; margin-top:.4em; }
        .msg { font-size:.62em; color:#374649; margin-top:.4em; }
        </style>
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressPackageStartupMessages({ library(readr); library(dplyr); library(knitr); library(tibble) })
FIG <- function(file, width = "90%") {
  p <- file.path("..", "output", "figures", file)
  if (file.exists(p))
    cat(sprintf("![](%s){fig-align=\"center\" width=%s}\n", p, width))
  else
    cat("::: {.placeholder}\n<span class='ph-tag'>\u25a2 pending data</span><br>Figure pending: run the corresponding script (see README) to generate it.\n:::\n")
}
.rdc <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), col_types = readr::cols(.default = "c")), error = function(e) NULL)
.sc <- .rdc("projections/bcr_scenarios_table.csv")
BCR_C <- if (!is.null(.sc)) .sc$BCR[grepl("^Central", .sc$Scenario)] else "1.0"
BCR_SK <- if (!is.null(.sc)) .sc$BCR[grepl("skills earnings premium", .sc$Scenario)] else "2.5"
.as <- .rdc("projections/bcr_assumptions.csv")
EMP_C <- if (!is.null(.as)) sprintf("%.0f%%", 100*as.numeric(.as$Central[.as$Parameter == "Female labour force participation"])) else "52%"
.mc <- .rdc("models/mincer_ifls_controls.csv")
MINCER_C <- if (!is.null(.mc)) { v <- .mc$`Return per year`[.mc$Sex == "Women" & grepl("^\\(4\\)", .mc$Model)]; sprintf("%.1f%%", 100 * as.numeric(v[1])) } else "9.6%"
.tn <- .rdc("projections/bcr_tornado.csv")
if (!is.null(.tn)) { .o <- .tn[!grepl("NEET|premium", .tn$Parameter), ]; .v <- as.numeric(c(.o$`BCR at low value`, .o$`BCR at high value`))
  RANGE_LO <- sprintf("%.2f", min(.v)); RANGE_HI <- sprintf("%.2f", max(.v)) } else { RANGE_LO <- "0.71"; RANGE_HI <- "1.57" }
TAB <- file.path("..", "output", "tables", "t8_target_funnel.csv")
```

## Contents

::: {.toc-grid}
**1 · The problem**: who is left out, where, and at what cost

**2 · Theory of change**: what the programme does and why it should work

**3 · Return on investment**: key assumptions and results

**4 · Key messages**

**Appendix**: regression results, methods and data
:::

# 1 · The problem {.section-divider background-color="#00377C"}

## Girls have the same school attendance rate as boys, but their NEET rate is close to double {.smaller}
```{r f1a, results='asis'}
FIG("f1a_outcomes_by_sex_en.png", "78%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
NEET: not in employment, education or training. Girls 15–24: 51% in school (boys 50%), 23% NEET (boys 13%), 25% employed (boys 37%). The gap opens at the transition out of school. SUSENAS Maret 2025, survey-weighted, 95% CIs.
:::

## East Java's NEET gender gap is wider than the national average {.smaller}
```{r f19, results='asis'}
FIG("f19_national_benchmark_en.png", "76%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls-minus-boys NEET gap: 10.3 points in East Java against 7.9 points nationally. Same survey, same definition, all provinces.
:::

## The gap costs an estimated USD 4.1 billion in forgone lifetime productivity {.smaller}
```{r f15, results='asis'}
FIG("f15_lifetime_loss_en.png", "80%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
About 288,000 girls are NEET in excess of the boys' rate. Each would be expected to earn about USD 14,100 more over a working lifetime (present value) if she had the boys' NEET rate.
:::

## Madura has higher NEET for girls and boys, with a similar gender gap {.smaller}
```{r f20, results='asis'}
FIG("f20_madura_en.png", "82%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls' NEET is 34% in Madura (Bangkalan, Sampang, Pamekasan, Sumenep) against 22% in the rest of East Java; boys' NEET is 23% against 12%. The gender gap is similar (11 vs 10 points), and fewer girls are in school (45% vs 52%). Madura is a priority area for a pilot.
:::

## Child marriage and outcomes: rest of East Java {.smaller}
```{r f24b, results='asis'}
FIG("f24_marriage_rest_en.png", "76%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 15–24 outside Madura, by whether they married before 18, survey-weighted (95% CIs). Regression tests by region are in the appendix. Early pregnancy before 18 overlaps almost entirely with child marriage in this data, so only child marriage is shown. Association, not causation.
:::

## Child marriage and outcomes: Madura {.smaller}
```{r f24a, results='asis'}
FIG("f24_marriage_madura_en.png", "76%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 15–24 in Bangkalan, Sampang, Pamekasan and Sumenep. Fewer girls, so wider intervals. Association, not causation.
:::

## Married girls who work do more unpaid family work and less wage employment {.smaller}
```{r f21, results='asis'}
FIG("f21_employment_type_en.png", "72%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Among employed girls married before 18, 35% are wage employees (others 75%) and 31% are unpaid family workers (others 12%); 22% work under 20 hours a week (others 12%). Based on 78 employed married girls, so indicative. SUSENAS has no earnings; the appendix shows earnings by type of work for young women from IFLS5.
:::

## Girls' NEET fell from 2024 to 2025, slightly more than in the rest of Indonesia {.smaller}
```{r f23, results='asis'}
FIG("f23_trend_ej_vs_national_en.png", "70%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls' NEET fell 5.6 points in East Java and 4.5 in the rest of Indonesia; child marriage fell 1.3 and 0.7 points. Tertiary enrolment at 18–20 rose 3.2 points, less than elsewhere (5.0). Both years use the same definition (any work in the last week, or temporarily absent from a job; not attending school). Most of the decline is national, so it cannot be attributed to a programme. A district-level comparison with programme districts is pending.
:::

## Senior-secondary enrolment also rose, in East Java and nationally {.smaller}
```{r f25, results='asis'}
FIG("f25_secondary_enrolment_trend_en.png", "70%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 16–18 enrolled in senior secondary, 2024 and 2025. A rise shared with the rest of Indonesia points to national drivers (such as PIP) rather than something specific to East Java.
:::

# 2 · Theory of change {.section-divider background-color="#00377C"}

## Completing senior secondary is associated with 20 points lower NEET for girls {.smaller}
```{r f22, results='asis'}
FIG("f22_completion_left_school_en.png", "74%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Among 18–20 year olds who have left school, girls who completed senior secondary have a NEET rate of 41%, against 62% for those who did not. Adjusted for age and area (girls 18–24), completion is associated with a 20-point lower NEET rate (p&lt;0.001). For boys there is no observed difference. Keeping girls in school to completion is the programme's main lever.
:::

## Evidence from Indonesia {.smaller}

::: {.tight-table}
| Programme | Design | Result | Relevance |
|---|---|---|---|
| **PKH** cash transfer (Priebe &amp; Sumarto, 2025) | Quasi-experimental | Child marriage −3.5pp among girls | A cash component reduces child marriage |
| **PIP** education assistance (Ulfa &amp; Rezki, 2024) | Matching and IV | School participation +11.4pp (pooled, not girls-specific) | Anchors the effect used in the ROI, with a transferability assumption |
| **PIP**, girls' senior secondary (Caniago et al., 2021) | Observational regression | Girls' senior-secondary enrolment +24.6pp (adjusted association) | Targets the stage where girls drop out |
| **Vocational vs general secondary schooling** (Pritadrajati, 2022) | Quasi-experimental | Female LFPR +3.4pp (significant at 10%); income +5.6% (not significant) | Used only as an uncertain scenario in the ROI |
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
In SUSENAS 2025, girls in PKH households have lower child-marriage rates than other girls (3% vs 4–6%), even though PKH targets poorer households. Consistent with the PKH evidence, though not a causal estimate.
:::

## Theory of change

```{mermaid}
%%| fig-width: 13
flowchart LR
  T["Target: girls 15–18<br/>poorest households,<br/>Madura districts"]:::t
  A1["Cash+: link to PKH / PIP,<br/>mentoring and<br/>marriage-delay outreach"]:::a
  A2["Digital and 21st-century<br/>skills in Double Track"]:::a
  O1["Child marriage<br/>delayed"]:::o
  O2["Girls stay in school<br/>and complete<br/>senior secondary"]:::o
  O3["Skills for<br/>work"]:::o
  R["Transition to work<br/>or higher education:<br/>lower NEET"]:::r
  I["Higher lifetime<br/>earnings"]:::r
  T --> A1
  T --> A2
  A1 --> O1
  A1 --> O2
  A2 --> O3
  O1 --> O2
  O2 --> R
  O3 --> R
  R --> I
  classDef t fill:#f4f4f4,stroke:#999,color:#333
  classDef a fill:#eaf6fc,stroke:#1CABE2,color:#00377C
  classDef o fill:#fff7e6,stroke:#FFC20E,color:#7a5b00
  classDef r fill:#e8f5e9,stroke:#2e7d32,color:#1b5e20
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Each arrow rests on evidence from Indonesia: cash transfers reduce girls' child marriage (PKH); education assistance raises school participation (PIP, pooled) and girls' senior-secondary enrolment (PIP, adjusted association); completion is associated with 20 points lower NEET for girls (this analysis). The link from skills to earnings is the weakest: the Indonesian estimate is not statistically significant. Mentoring and outreach have only international evidence for girls. The cash component links girls to existing PKH and PIP rather than creating a new transfer.
:::

# 3 · Return on investment {.section-divider background-color="#00377C"}

## Key assumptions {.smaller}

::: {.big-table}
| Assumption | Value | Source |
|---|---|---|
| Reduction in girls' NEET | 5 points | PIP attendance effect (+11.4pp), 45% assumed to come from girls who would be NEET |
| Extra schooling for a girl kept in school | 3 years | Senior-secondary cycle |
| Earnings gain per year of schooling | `r MINCER_C` | IFLS5 wage return for women, controlling for region, parents' education and selection into paid work |
| Female earnings | Rp20 million/year | Below East Java minimum wage 2025, reflecting informal work |
| Probability of earning (employment rate, women 25–54, East Java) | `r EMP_C` | SUSENAS Maret 2025 |
| Earnings growth / discount rate | 2% / 3% per year | Career growth; 3% real, with 2% and 5% in sensitivity |
| Skills earnings premium (scenario only, not in the central case) | 5.6% for 10 years | Point estimate for vocational vs general schooling (Pritadrajati 2022), not statistically significant |
| Programme cost | USD 200 per girl | USD 58 per year for 3 years, plus 15% overhead; Cash+ links to existing PKH/PIP |
:::

## How the return is built {.smaller}

::: {.vmid}
::: {.big-table}
```{r roi-steps}
st <- tryCatch(readr::read_csv(file.path("..","output","projections","bcr_steps.csv"), col_types = readr::cols(.default = "c")), error = function(e) NULL)
if (!is.null(st)) knitr::kable(st, align = "lr")
```
:::
:::

## Every USD 1 invested returns about USD `r BCR_C` in lifetime earnings {.smaller}

```{r f26, results='asis'}
FIG("f26_roi_bars_en.png", "60%")
```

::: {.kf-list style="font-size:.74em;"}
- **For every USD 1 invested, about USD `r BCR_C` comes back as lifetime earnings** of girls kept in school through senior secondary. A benefit-cost ratio, not a fiscal return to government; it counts earnings only and excludes delayed marriage, health and intergenerational benefits.
- **If the skills component also raised earnings by 5.6%** for every girl who works, the return would be about USD `r BCR_SK`. That estimate comes from vocational schooling in Indonesia and is not statistically significant, so it is a scenario, not the central case.
- **The USD 200 cost per girl** covers three years of digital content, teacher training, mentoring, shared connectivity and marriage-delay outreach (USD 58 a year), plus 15% overhead. Girls are linked to existing PKH and PIP transfers, which are not an additional cost.
:::

## Benefit-cost ratio by scenario {.smaller}

::: {.vmid}
::: {.tight-table}
```{r bcr-table}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
sc <- tryCatch(readr::read_csv(file.path("..","output","projections/bcr_scenarios_table.csv"), col_types = readr::cols(.default = "c")), error=function(e) NULL)
if (!is.null(sc)) knitr::kable(sc |> dplyr::select(-`Cost per NEET case averted (USD)`), align = "lrrrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
The central case is the retention channel; the second row adds the uncertain skills premium; the others show earnings at the 2025 minimum wage, and the cost of adding a Cash+ stipend for the poorest 40%. With other assumptions varied one at a time, the central return stays between USD `r RANGE_LO` and USD `r RANGE_HI` (appendix).
:::

# 4 · Key messages {.section-divider background-color="#00377C"}

## Key messages {.smaller}

::: {.kf-list style="font-size:.8em;"}
**1.** Girls in East Java have the same school attendance rate as boys but are close to twice as likely to be NEET, a wider gap than nationally, costing about USD 4.1 billion in forgone lifetime productivity.

**2.** Completing senior secondary is associated with 20 points lower NEET for girls; child marriage with leaving school and more unpaid family work. Both are concentrated among poorer girls and in Madura.

**3.** A USD 200-per-girl package returns about USD `r BCR_C` per USD 1 in lifetime earnings from keeping girls in school, before any benefit from skills, delayed marriage or health.
:::

## Recommendations {.smaller}

::: {.kf-list style="font-size:.78em;"}
**1. Target senior-secondary completion, not enrolment.** Girls already enrol at the same rate as boys; the loss happens at the transition. Set programme targets on completion and the first job, measured through SUSENAS.

**2. Start in Madura and the poorest 40% of households.** NEET is highest and attendance lowest there, and child marriage falls steeply with household wealth, so each dollar does most there.

**3. Link girls to PKH and PIP rather than creating a new transfer.** Indonesian evidence shows PKH reduces girls' child marriage; linkage uses existing systems. Cost a new stipend only if coverage beyond current entitlements is wanted.

**4. Fund pedagogy and mentoring, not devices.** Internet use is already near universal, and devices alone did not raise learning in Uruguay.

**5. Measure the programme's own effect.** The 2024-2025 fall in girls' NEET is national, not provincial. Compare programme and non-programme districts, and track whether girls move into wage employment rather than unpaid family work.
:::

# Appendix {.section-divider background-color="#00377C"}

## Regression results: child marriage and outcomes {.smaller}

::: {.vmid}
::: {.tight-table}
```{r marriage-tests}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
mt <- rd("models/marriage_tests.csv")
if (!is.null(mt)) knitr::kable(mt |> dplyr::select(Hypothesis, `Effect (pp)`, SE, p, N), align = "lrrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Linear-probability models, girls 15–24, survey-weighted, PSU-clustered. **Controls: age and urban/rural.** Poverty is not controlled for because the 2025 expenditure module is not yet available. The association with employment is small and not statistically significant. Association, not causation.
:::

## Child marriage and outcomes by region: regression results {.smaller}

::: {.vmid}
::: {.tight-table}
```{r marriage-region-tests}
mr <- tryCatch(readr::read_csv(file.path("..","output","models","followup2_A_marriage_tests_by_region.csv"), show_col_types = FALSE), error = function(e) NULL)
if (!is.null(mr)) knitr::kable(mr, align = "llrrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Linear-probability models, girls 15–24, survey-weighted, PSU-clustered; controls for age and urban/rural. Effects in percentage points.
:::

## Earnings by type of work, young women (IFLS5) {.smaller}

::: {.vmid}
::: {.tight-table}
```{r earnings-status}
es <- tryCatch(readr::read_csv(file.path("..","output","models","followup2_C_earnings_by_status_ifls.csv"), col_types = readr::cols(.default = "c")), error = function(e) NULL)
if (!is.null(es)) knitr::kable(es, align = "lrrr")
```
:::
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Women 15–30, IFLS5 (2014/15 prices). SUSENAS collects no earnings. Unpaid family workers have no earnings by definition.
:::

## NEET among girls, by district {.smaller}
```{r f6, results='asis'}
FIG("f6_neet_by_kabkota_en.png", "60%")
```


## Digital access and use {.smaller}
```{r f2, results='asis'}
FIG("f2_digital_by_area_sex_en.png", "86%")
```


## Early pregnancy and outcomes {.smaller}
```{r f13, results='asis'}
FIG("f13_pregnancy_pathway_en.png", "78%")
```

::: {style="font-size:0.58em; color:#7a7a7a;"}
Recorded only for ever-married women, so a lower bound. Girls with a first pregnancy before 18 are almost all girls married before 18, so the main deck reports child marriage only.
:::

## Child marriage and NEET by household PKH receipt {.smaller}

::: {.tight-table}
```{r pkh}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
pk <- rd("models/followup_D_pkh.csv")
if (!is.null(pk)) knitr::kable(pk, align = "llrrr")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Girls 15–24, SUSENAS 2025. PKH targets poorer households, so lower rates among recipients are consistent with a protective effect, but are not a causal estimate.
:::

## Cost components {.smaller}

::: {.big-table}
```{r costing-table}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
ct <- rd("projections/bcr_costing_table.csv")
if (!is.null(ct)) knitr::kable(ct, align = "lrl")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
These are benchmark unit costs from comparable programmes, not quotes from Double Track; they are placeholders until the provincial government shares actual budgets. The transfer itself is not costed because the Cash+ component links girls to existing PKH and PIP entitlements; a new stipend for the poorest 40% at the PIP senior-secondary rate would add about USD 46 per girl per year (ROI table).
:::

## Sensitivity of the return to each assumption {.smaller}

::: {.tight-table}
```{r tornado}
rd <- function(p) tryCatch(readr::read_csv(file.path("..","output",p), show_col_types=FALSE), error=function(e) NULL)
tn <- rd("projections/bcr_tornado.csv")
if (!is.null(tn)) knitr::kable(tn |> dplyr::select(-spread), align = "lrr")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
Each assumption varied on its own, others held at central values.
:::

::: {.kf-list style="font-size:.7em; margin-top:.6em;"}
- **Low values**: a 2.5-point NEET effect (half the PIP-based estimate), the consumption-based return of 5.6% instead of 8%, earnings of Rp15m instead of Rp20m, a 5% discount rate, or a cost of USD 75 a year instead of 58.
- **High values**: a 7.5-point NEET effect, the 12.2% return without the selection correction, earnings of Rp25m, a 2% discount rate, or a cost of USD 45 a year.
- The effect size and the return to schooling matter most; no single assumption on its own takes the central case far from break-even.
:::

## What the return counts {.smaller}

::: {.bcr-grid}
<div class="bcr-col counted">
<h4>Counted</h4>
Higher schooling leads to a lifetime earnings premium for girls moved out of NEET, using the female return to schooling over a discounted working life.
</div>
<div class="bcr-col notcounted">
<h4>Excluded <span style="font-weight:400;">(so the return is a lower bound)</span></h4>
Delayed child marriage · health and productivity spillovers · intergenerational gains · fiscal returns · broader empowerment
</div>
:::

## Comparable international evidence {.smaller}

```{r evidence-table}
tibble::tribble(
  ~Study, ~Design, ~Sample, ~Finding, ~"Relevance for East Java",
  "Bergstrom &amp; Özler 2021", "Review of 104 interventions from 70 studies", "Adolescent girls", "Mixed evidence for life-skills training and girls' clubs; stronger for cash and schooling support", "Supports cash and schooling components; cautions on clubs alone",
  "Bandiera et al. 2018 (Sierra Leone)", "RCT", "Girls 12–25", "Enrolment +8.5pp during the Ebola period; effects on income-generating activity vary by age", "Girls' clubs protected schooling in a crisis context",
  "Aramburu, Goicoechea &amp; Mobarak 2021 (Argentina, Colombia)", "RCT", "Women 18–50", "Coding bootcamps raised coding skills", "Digital-skills content",
  "Näslund-Hadley et al. 2025 (Paraguay)", "RCT", "Second-grade pupils", "+0.09 SD computational thinking", "Pedagogy; limited transfer to adolescent employment",
  "Gülen et al. 2025 (Turkey)", "Nonrandomized controlled pre/post", "Girls 14–16", "Higher STEM aspirations and motivation", "Girls' STEM engagement",
  "Prakash et al. 2019, Samata (India)", "RCT", "Poor girls 13–16", "No overall effect on completion or marriage; schooling gains in one of two districts", "Transition support alone had limited effects",
  "Edmonds et al. 2021, Room to Read (India)", "RCT", "Girls", "Dropout 4.1pp lower (main survey estimate)", "Mentoring and life skills reduce dropout",
  "Ashraf et al. 2025, Ten Years of Relational Power (Zambia)", "RCT, 10-year follow-up", "Girls ~14", "+0.23 years of schooling", "Durable, low-cost soft-skills layer",
  "de Melo et al. 2014, Plan Ceibal (Uruguay)", "Quasi-experimental", "Pupils", "No detectable maths or reading gains in the first two years", "Devices alone did not raise learning; fund pedagogy",
  "Faizah, Isbah &amp; Azca 2025 (online; 2026 issue)", "Qualitative", "Indonesian youth", "Norms, care duties and marriage drive girls' NEET", "Indonesia-specific NEET context"
) |> knitr::kable(align = "lllll")
```

::: {style="font-size:0.55em; color:#7a7a7a;"}
Findings as reported in each source, with design and sample as verified. Used to inform programme design; the ROI effect comes from the Indonesian evidence. References on the last slide.
:::

## Returns to schooling: specification and results {.smaller}

::: {style="font-size:0.8em"}
**Wage-based** (IFLS5 2014/15, adults 20–60, estimated separately for women and men, weighted):

$$\ln(w_i) = \beta_0 + \beta_1 S_i + \beta_2 \text{Exp}_i + \beta_3 \text{Exp}_i^2 + \beta_4 \text{Urban}_i + \gamma_p + \phi_f + \phi_m + \delta \lambda_i + \varepsilon_i$$

$w$: monthly wage; $S$: years of schooling; Exp: potential experience (age minus schooling minus 6); $\gamma_p$: province fixed effects; $\phi_f$, $\phi_m$: father's and mother's education (fixed effects); $\lambda$: selection term from a first-stage model of working, in which children under 5 in the household predict whether a woman works but not her wage (Heckman two-step).
:::

::: {.tight-table}
```{r mincer-controls}
mc <- tryCatch(readr::read_csv(file.path("..","output","models","mincer_ifls_controls.csv"), show_col_types = FALSE), error = function(e) NULL)
if (!is.null(mc)) knitr::kable(mc |> dplyr::mutate(`Return per year` = sprintf("%.1f%%", 100 * `Return per year`),
                                                   SE = sprintf("%.1f", 100 * SE)) |>
                                 dplyr::select(Sex, Model, `Return per year`, SE, N), align = "llrrr")
```
:::

::: {style="font-size:0.58em; color:#7a7a7a;"}
The ROI uses the women's estimate with all controls (model 4, `r MINCER_C`). The consumption-based return from SUSENAS 2024 (5.6% for women; household per-capita expenditure regressed on schooling, schooling², age, age² and urban) is the low case in the sensitivity analysis.
:::

## Data sources {.smaller}

::: {style="font-size:0.95em"}
- **SUSENAS Maret 2025** (BPS): NEET, school attendance and completion, child marriage, early pregnancy, type of work and hours, PKH receipt and digital access, for East Java and all provinces. East Java: 98,738 individuals, 12,599 aged 15–24, of whom 6,060 girls.
- **SUSENAS Maret 2024** (BPS): the 2024 comparison, and per-capita expenditure for the consumption-based return to schooling.
- **IFLS5 2014/15**: wages for the wage-based return to schooling.
- **Sakernas** (BPS): female labour force participation in East Java.
- **UMP Jawa Timur 2025**: benchmark for female earnings.
- **Published evaluations** of PKH, PIP and vocational training: the programme effect used in the return on investment.

All SUSENAS estimates are survey-weighted, with PSU/strata-clustered standard errors.
:::

## Definitions {.smaller}

::: {style="font-size:0.95em"}
- **NEET (15–24)**: not in employment, education or training, built as not attending school (`R611`; 2024: `R610`) and not working in the last week or temporarily absent from a job (`R703_A`; `R704`, 2024: `R705`).
- **Married before 18**: age at first marriage (`R409`) under 18.
- **Senior-secondary completion**: highest certificate (`R615`) at SMA/SMK/MA level or above.
- **Type of work**: status in main job (`R706`); hours (`R707`).
- **Madura**: Bangkalan, Sampang, Pamekasan, Sumenep.

Several variables were renumbered between the 2024 and 2025 questionnaires (school attendance R610 to R611, temporarily absent R705 to R704, certificate R614 to R615, employment status R707 to R706); the activity block (R703) is the same in both years and both years are coded identically. The operational NEET definition does not separately exclude participants in non-formal training, which SUSENAS does not record.
:::

## References {.smaller}

::: {style="font-size:0.5em; line-height:1.35"}
Aramburu, Goicoechea and Mobarak (2021). Coding Bootcamps for Female Digital Employment: Evidence from an RCT in Argentina and Colombia. World Bank. · Ashraf et al. (2025). Ten Years of Relational Power. · Bandiera et al. (2018). The Economic Lives of Young Women in the Time of Ebola. · Bergstrom and Özler (2021). Improving the Well-Being of Adolescent Girls in Developing Countries. World Bank. · Caniago et al. (2021), Table 5. · de Melo, Machado and Miranda (2014). The Impact of a One Laptop per Child Program on Learning: Evidence from Uruguay. · Edmonds, Feigenberg and Leight (2021). Advancing the Agency of Adolescent Girls. · Faizah, Isbah and Azca (2025, online; 2026 issue). · Gülen et al. (2025). · Näslund-Hadley et al. (2025). IDB. · Prakash et al. (2019). Samata. · Priebe and Sumarto (2025). PKH and child marriage. · Pritadrajati (2022). Bulletin of Monetary Economics and Banking. · Ulfa and Rezki (2024), Table 4. · BPS, SUSENAS Maret 2024 and 2025; IFLS5 (2014/15); UMP Jawa Timur 2025.

Links and table or page references to be completed for each study before circulation.
:::
